---
title: "Fantasy Football 2018"
author: "Conner Tolley"
date: "October 8, 2018"
output: 
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

# Introduction

This is a practice markdown notebook to analyze fantasy football data for the 2018 season. The data can be found here https://www.pro-football-reference.com/years/2018/fantasy.htm. 

There are three parts to my script as follows:

* Feature engineering
* Missing value imputation
* Prediction!

## Load and check data

```{r, message=FALSE}
# Load Packages

library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```
Now that the packages are loaded, let's read in and take a peek at the data.


```{r, message=FALSE, warning=FALSE}
# Read the csv file. Skip the first row and use second row as headers for the data frame
full <- read.csv('FantasyRankings2018.csv', stringsAsFactors = F, skip = 1)

# Rename columns to more readable names
full <- rename(full, Name = X, Team = Tm, Position = FantPos, Num = Rk, Games_Played = G, Games_Started = GS, Completed_Passing = Cmp, Attempts_Passing = Att, Yards_Passing = Yds, TD_Passing = TD, Ints_Passing = Int, Attempts_Rushing = Att.1, Yards_Rushing = Yds.1, Yards_Attempt_Rushing = Y.A, TD_Rushing = TD.1, Tgt_Rec = Tgt, Receptions = Rec, Yards_Rec = Yds.2, Yards_Per_Rec = Y.R, TD_Rec = TD.2, X2P_Made = X2PM, X2P_Passes = X2PP, Draft_Kings = DKPt, Fan_Duel = FDPt)

# check data
 str(full)
 
 # Remove the nickname from the end of the Name column after the "\\"
full$Name <- gsub("\\\\.*", "", full$Name)
```

## Teams

Let's take a look at how the teams are doing with fantasy points.

```{r}
# Change Josh Gordon's team to New England


# Filter out the players with no fantasy points
full <- filter(full, PPR != "", FantPt != "", Team != '2TM')

# Group by number of points per team and positions
full <- 
  full %>%
  filter(Position != "") %>%
  group_by(Team, Position) %>%
  summarise(Total_Points = sum(PPR)) 

# Rename team names to full name of team
full$Team[which(full$Team == "ARI")] <- "Arizona"
full$Team[which(full$Team == "BUF")] <- "Buffalo"
full$Team[which(full$Team == "TEN")] <- "Tennessee"
full$Team[which(full$Team == "WAS")] <- "Washington"
full$Team[which(full$Team == "DAL")] <- "Dallas"
full$Team[which(full$Team == "CHI")] <- "Chicago"
full$Team[which(full$Team == "CAR")] <- "Carolina"
full$Team[which(full$Team == "NYJ")] <- "New York Jets"
full$Team[which(full$Team == "MIA")] <- "Miami"
full$Team[which(full$Team == "SEA")] <- "Seattle"
full$Team[which(full$Team == "CLE")] <- "Cleveland"
full$Team[which(full$Team == "TAM")] <- "Tampa Bay"
full$Team[which(full$Team == "DEN")] <- "Denver"
full$Team[which(full$Team == "NYG")] <- "New York Giants"
full$Team[which(full$Team == "SFO")] <- "San Francisco"
full$Team[which(full$Team == "NWE")] <- "New England"
full$Team[which(full$Team == "PHI")] <- "Philadelphia"
full$Team[which(full$Team == "JAX")] <- "Jacksonville"
full$Team[which(full$Team == "DET")] <- "Detroit"
full$Team[which(full$Team == "HOU")] <- "Houston"
full$Team[which(full$Team == "OAK")] <- "Oakland"
full$Team[which(full$Team == "CIN")] <- "Cincinnati"
full$Team[which(full$Team == "BAL")] <- "Baltimore"
full$Team[which(full$Team == "GNB")] <- "Green Bay"
full$Team[which(full$Team == "IND")] <- "Indinapolis"
full$Team[which(full$Team == "MIN")] <- "Minnesota"
full$Team[which(full$Team == "LAC")] <- "Los Angeles Chargers"
full$Team[which(full$Team == "ATL")] <- "Atlanta"
full$Team[which(full$Team == "PIT")] <- "Pittsburg"
full$Team[which(full$Team == "KAN")] <- "Kansas City"
full$Team[which(full$Team == "LAR")] <- "Los Angeles Rams"
full$Team[which(full$Team == "NOR")] <- "New Orleans"

# Plot into a bar graph
ggplot(data = full, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = "Team Points after Week 5 (ppr)") +
  coord_flip() +
  theme_few()

```

Now We can look at some plots based on position and team.

```{r}
# Filter to just get the QBs
fullQB <- 
  full %>%
  filter(Position == "QB")

# Plot QBs into a bar graph
ggplot(data = fullQB, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = "Quarterback Points after Week 5 (ppr)") +
  coord_flip() +
  theme_few()

# Filter to just get the TEs
fullTE <- 
  full %>%
  filter(Position == "TE")

# Plot QBs into a bar graph
ggplot(data = fullTE, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = "Tight End Points after Week 5 (ppr)") +
  coord_flip() +
  theme_few()

# Filter to just get the RBs
fullRB <- 
  full %>%
  filter(Position == "RB")

# Plot QBs into a bar graph
ggplot(data = fullRB, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = "Running Back Points after Week 5 (ppr)") +
  coord_flip() +
  theme_few()

# Filter to just get the WRs
fullWR <- 
  full %>%
  filter(Position == "WR")

# Plot QBs into a bar graph
ggplot(data = fullWR, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = "Wide Receiver Points after Week 5 (ppr)") +
  coord_flip() +
  theme_few()

```
