---
title: "Fantasy Football 2018"
author: "Conner Tolley"
date: "October 8, 2018"
output: 
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

# Introduction

This is a practice markdown notebook to analyze fantasy football data for the 2018 season. The data can be found here https://www.pro-football-reference.com/years/2008/fantasy.htm. 

There are three parts to my script as follows:

* Feature engineering
* Missing value imputation
* Prediction!

## Load and check data

```{r, message=FALSE}
# Load Packages

library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
```
Now that the packages are loaded, let's read in and take a peek at the data.


```{r, message=FALSE, warning=FALSE}
# Make the year a variable we can change
YEAR <- "2018"

# Get the file name based on the year we supply
fileName <- sprintf("Data_Files/fantData%s.csv", YEAR)

# Read the csv file
full <- read.csv(fileName, stringsAsFactors = F)

# Rename columns to more readable names
full <- rename(full, Name = X, Team = Tm, Position = FantPos, Num = Rk, Games_Played = G, Games_Started = GS, Completed_Passing = Cmp, Attempts_Passing = Att, Yards_Passing = Yds, TD_Passing = TD, Ints_Passing = Int, Attempts_Rushing = Att.1, Yards_Rushing = Yds.1, Yards_Attempt_Rushing = Y.A, TD_Rushing = TD.1, Tgt_Rec = Tgt, Receptions = Rec, Yards_Rec = Yds.2, Yards_Per_Rec = Y.R, TD_Rec = TD.2, X2P_Made = X2PM, X2P_Passes = X2PP, Draft_Kings = DKPt, Fan_Duel = FDPt)

# check data
 str(full)
 
 # Remove the nickname from the end of the Name column after the "\\"
full$Name <- gsub("\\\\.*", "", full$Name)
```

## Teams

Let's take a look at how the teams are doing with fantasy points.

```{r}
# Change Josh Gordon's team to New England
full$Team[full$Name == "Josh Gordon"] <- "NWE"

# Filter out the players with no fantasy points
full <- filter(full, PPR != "", FantPt != "", Team != '2TM', Team != '3TM')

# Rename team names to full name of team
full$Team[which(full$Team == "ARI")] <- "Arizona"
full$Team[which(full$Team == "BUF")] <- "Buffalo"
full$Team[which(full$Team == "TEN")] <- "Tennessee"
full$Team[which(full$Team == "WAS")] <- "Washington"
full$Team[which(full$Team == "DAL")] <- "Dallas"
full$Team[which(full$Team == "CHI")] <- "Chicago"
full$Team[which(full$Team == "CAR")] <- "Carolina"
full$Team[which(full$Team == "NYJ")] <- "New York Jets"
full$Team[which(full$Team == "MIA")] <- "Miami"
full$Team[which(full$Team == "SEA")] <- "Seattle"
full$Team[which(full$Team == "CLE")] <- "Cleveland"
full$Team[which(full$Team == "TAM")] <- "Tampa Bay"
full$Team[which(full$Team == "DEN")] <- "Denver"
full$Team[which(full$Team == "NYG")] <- "New York Giants"
full$Team[which(full$Team == "SFO")] <- "San Francisco"
full$Team[which(full$Team == "NWE")] <- "New England"
full$Team[which(full$Team == "PHI")] <- "Philadelphia"
full$Team[which(full$Team == "JAX")] <- "Jacksonville"
full$Team[which(full$Team == "DET")] <- "Detroit"
full$Team[which(full$Team == "HOU")] <- "Houston"
full$Team[which(full$Team == "OAK")] <- "Oakland"
full$Team[which(full$Team == "CIN")] <- "Cincinnati"
full$Team[which(full$Team == "BAL")] <- "Baltimore"
full$Team[which(full$Team == "GNB")] <- "Green Bay"
full$Team[which(full$Team == "IND")] <- "Indianapolis"
full$Team[which(full$Team == "MIN")] <- "Minnesota"
full$Team[which(full$Team == "LAC")] <- "Los Angeles Chargers"
full$Team[which(full$Team == "ATL")] <- "Atlanta"
full$Team[which(full$Team == "PIT")] <- "Pittsburgh"
full$Team[which(full$Team == "KAN")] <- "Kansas City"
full$Team[which(full$Team == "LAR")] <- "Los Angeles Rams"
full$Team[which(full$Team == "NOR")] <- "New Orleans"
full$Team[which(full$Team == "SDG")] <- "San Diego"
full$Team[which(full$Team == "STL")] <- "St. Louis"

# Add an average column
full$AveragePPR <- full$PPR / full$Games_Played

print(head(full))

# # Full with names
# # Group by number of points per team and positions
# fullWithNames <- 
#   full %>%
#   filter(Position != "") %>%
#   group_by(Team, Position, Name) %>%
#   summarise(Total_Points = sum(PPR)) 
# 
# # Filter to just get the RBs
# fullRBWithNames <- 
#   fullWithNames %>%
#   filter(Position == "RB")
# 
# # Plot QBs into a bar graph
# ggplot(data = fullRBWithNames, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Name)) +
#   geom_bar(stat = "identity") +
#   guides(fill = FALSE) +
#   coord_flip() +
#   labs(x = "Teams") +
#   labs(title = sprintf("Running Back Points for %s (ppr)", YEAR)) +
#   theme_few()

# Group by number of points per team and positions
fullAverage <- 
  full %>%
  filter(Position != "") %>%
  group_by(Team, Position) %>%
  summarise(Total_Points = sum(AveragePPR)) 

# Group by number of points per team and positions
full <- 
  full %>%
  filter(Position != "") %>%
  group_by(Team, Position) %>%
  summarise(Total_Points = sum(PPR)) 

# Plot into a bar graph for total team points
ggplot(data = full, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Total Team Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

# Plot into a bar graph for average team points per game
ggplot(data = fullAverage, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Average Team Points per game for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

```

Now We can look at some plots based on position and team.

```{r}
# Filter to just get the QBs
fullQB <- 
  full %>%
  filter(Position == "QB")

# Plot QBs into a bar graph
ggplot(data = fullQB, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Quarterback Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

# Filter to just get the TEs
fullTE <- 
  full %>%
  filter(Position == "TE")

# Plot QBs into a bar graph
ggplot(data = fullTE, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Tight End Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

# Filter to just get the RBs
fullRB <- 
  full %>%
  filter(Position == "RB")

# Plot QBs into a bar graph
ggplot(data = fullRB, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Running Back Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

# Filter to just get the WRs
fullWR <- 
  full %>%
  filter(Position == "WR")

# Plot QBs into a bar graph
ggplot(data = fullWR, aes(x=reorder(Team, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Wide Receiver Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

```

# Divisions
Lets take a look at how the divisions are scoring in fantasy. We'll need to create a new column for divisions and assign a division to each team based on the team name.

```{r}
# Add a division column
full$Division <- NA

# Assign teams to the AFC East division
full$Division[full$Team == "New England" |
                full$Team == "Buffalo" |
                full$Team == "Miami" | 
                full$Team == "New York Jets"] <- "AFC East"

# Assign teams to the AFC North division
full$Division[full$Team == "Pittsburgh" |
                full$Team == "Baltimore" |
                full$Team == "Cincinnati" | 
                full$Team == "Cleveland"] <- "AFC North"

# Assign teams to the AFC South division
full$Division[full$Team == "Tennessee" |
                full$Team == "Indianapolis" |
                full$Team == "Houston" | 
                full$Team == "Jacksonville"] <- "AFC South"

# Assign teams to the AFC West division
full$Division[full$Team == "Kansas City" |
                full$Team == "Los Angeles Chargers" |
                full$Team == "Oakland" | 
                full$Team == "Denver"] <- "AFC West"

# Assign teams to the NFC East division
full$Division[full$Team == "Philadelphia" |
                full$Team == "Dallas" |
                full$Team == "Washington" |
                full$Team == "New York Giants"] <- "NFC East"

# Assign teams to the NFC North division
full$Division[full$Team == "Minnesota" |
                full$Team == "Detroit" |
                full$Team == "Green Bay" |
                full$Team == "Chicago"] <- "NFC North"

# Assign teams to the NFC South division
full$Division[full$Team == "New Orleans" |
                full$Team == "Carolina" |
                full$Team == "Atlanta" |
                full$Team == "Tampa Bay"] <- "NFC South"

# Assign teams to the NFC West division
full$Division[full$Team == "Arizona" |
                full$Team == "Los Angeles Rams" |
                full$Team == "Seattle" |
                full$Team == "San Francisco"] <- "NFC West"

# Plot Divisions into a bar graph
ggplot(data = full, aes(x=reorder(Division, -Total_Points), y=Total_Points, fill=Position)) +
  geom_bar(stat = "identity") +
  labs(x = "Teams") +
  labs(title = sprintf("Division Points for %s (ppr)", YEAR)) +
  coord_flip() +
  theme_few()

```